---
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
    template: ~/workspace/cv/git/svm-latex-ms.tex
title: "Homework 5"
author: 
  name: "Joyce Yu Cahoon"
#abstract: "Application of sequential monte carlo (SMC) to the development of IM plausibility functions for the Nile."
# bibliography: /Users/jcahoon/workspace/metar/ref.bib
#bibliography: /home/joyce/workspace/metar/ref.bib
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
endnote: no
header-includes:
- \setlength{\abovedisplayskip}{.2pt}
- \setlength{\belowdisplayskip}{.2pt}
- \usepackage{placeins}
- \usepackage{setspace}
- \usepackage{chngcntr}
- \usepackage{multicol}
- \usepackage{lscape}
- \counterwithin{figure}{section}
- \counterwithin{table}{section}
- \usepackage{mathrsfs}
- \usepackage{mathtools}
- \usepackage{multirow}
- \newtheorem{theorem}{Theorem}
- \usepackage[linesnumbered,algoruled,boxed,lined,commentsnumbered]{algorithm2e} 
- \usepackage{bm}
- \usepackage{framed}
- \usepackage{xcolor}
- \let\oldquote=\quote
- \let\endoldquote=\endquote
- \colorlet{shadecolor}{orange!15}
- \renewenvironment{quote}{\begin{shaded*}}{\end{shaded*}}
- \newcommand{\minus}{\scalebox{0.75}[1.0]{$-$}}
- \newcommand{\V}[1]{{\bm{{#1}}}}
--- 

```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(quantmod)
library(RiskPortfolios)
```

# 7.7

Show that the solution to 7.18 is given by 7.17. As a further generalization, what is the solution to the following problem: minimizing with respect the symmetric matrix $\V{S}$: 

$$
||\hat{\V{\Sigma}}_{\lambda} - \V{S}||^2_{F} \quad \text{such that} \quad \lambda_{\min}(\V{S}) \geq \delta
$$

We can show the solution to 7.18 is given by 7.17 since:
$$
\begin{aligned}
||\hat{\V{\Sigma}}_{\lambda} - \V{S}||^2_{F} &= tr[(\hat{\V{\Sigma}}_{\lambda} - \V{S})(\hat{\V{\Sigma}}_{\lambda} - \V{S})^T] \\
&= tr(\hat{\V{\Sigma}}\hat{\V{\Sigma}}^T) - tr(\hat{\V{\Sigma}} \V{S}^T) - tr(\V{S} \hat{\V{\Sigma}}^T) + tr(\V{S}\V{S}^T) \\
&= tr(\hat{\V{\Sigma}} \hat{\V{\Sigma}}^T) - 2tr(\hat{\V{\Sigma}} \V{S}^T) +  tr(\V{S}\V{S}^T) 
\end{aligned}
$$
Since we eventually want to constrain $\V{S}$ to positive definite matrices, let's represent $\V{S}$ as $\V{V}\V{V}^T$, leading to:
$$
\begin{aligned}
tr(\hat{\V{\Sigma}} \hat{\V{\Sigma}}^T) - 2tr(\hat{\V{\Sigma}} \V{S}\V{S}^T) +  tr(\V{V}\V{V}^T\V{V}\V{V}^T) 
\end{aligned}
$$
Taking the derivative of this relative to $\V{V}$ results in $-4\V{V}^T\hat{\V{\Sigma}} + 4\V{V}^T\V{V}\V{V}^T = 0$. We can simplify and solve this by letting $\V{V} = \Gamma^T ({\Lambda}^+)^{1/2}$. Since our covariance matrix is symmetric and can be broken down into $\Gamma^T \Lambda \Gamma$, then: 
$$
\begin{gathered}
-4[(\Lambda^+)^{1/2} \Gamma \Gamma^T (\Lambda^+ + \Lambda^-)\Gamma] + 4(\Lambda^+)^{1/2}\Gamma \Gamma^T (\Lambda^+)^{1/2} (\Lambda^+)^{1/2} \Gamma \\ 
-4[(\Lambda^+)^{1/2} (\Lambda^+ + \Lambda^-) \Gamma] + 
4(\Lambda^+)^{1/2} \Lambda^+ \Gamma \\ 
 -4[(\Lambda^+)^{1/2}\Lambda^+ \Gamma ] + 4 (\Lambda^+)^{1/2} \Lambda^+\Gamma  
\end{gathered}
$$
Our positive definite matrice $\V{S} = \V{V}\V{V}^T$ thus minimizes the function, which can also be represented by 7.17 where:
$$
\hat{\V{\Sigma}}^+ = \V{\Lambda}^T diag(\lambda_1^+, \ldots, \lambda_p^+) \V{\Lambda}
$$

# 7.10

Let $\hat{\V{\Sigma}}$ be an estimated volatility matrix of a true volatility. Show that for any portfolio allocation $\V{w}$, the relative estimation error is bounded by:

$$
|\frac{\V{w}^T \hat{\V{\Sigma}} \V{w} }{ \V{w}^T \V{\Sigma} \V{w}}-1| \leq || \V{\Sigma} ^{-1/2} \hat{\V{\Sigma}} \V{\Sigma}^{-1/2} - \V{I}_p||
$$
As proven in 7.8: 
$$
tr[\hat{\V{\Sigma}} \V{\Sigma}^{-1} - \V{I}_p]^2 = || \V{\Sigma}^{-1/2} \hat{\V{\Sigma}} \V{\Sigma}^{-1/2} - \V{I}_p||_F^2
$$
Thus:
$$
tr[\hat{\V{\Sigma}} \V{\Sigma}^{-1} - \V{I}_p] = || \V{\Sigma}^{-1/2} \hat{\V{\Sigma}} \V{\Sigma}^{-1/2} - \V{I}_p||_F
$$
We can then simplify: 
$$
\begin{aligned}
|\frac{\V{w}^T \hat{\V{\Sigma}} \V{w} }{ \V{w}^T \V{\Sigma} \V{w}}-1| &=
|\frac{\V{w}^T(\hat{\V{\Sigma}} - \V{\Sigma}) \V{w}}{ \V{w}^T \V{\Sigma} \V{w} } |\\
&= || \frac{\hat{\V{\Sigma}} - \V{\Sigma} }{\V{\Sigma}}|| \\
&= ||\hat{\V{\Sigma}} \V{\Sigma}^{-1} - \V{I}_p || \\
&\leq ||\hat{\V{\Sigma}} \V{\Sigma}^{-1} - \V{I}_p ||_F \\
&\leq tr[\hat{\V{\Sigma}} \V{\Sigma}^{-1} - \V{I}_p] \\
&= || \V{\Sigma}^{-1/2} \hat{\V{\Sigma}} \V{\Sigma}^{-1/2} - \V{I}_p||_F
\end{aligned}
$$

# 7.11

Suppose that we have 100 investable stocks, labeled as 1 through 100 and classified as "Consumer Non-durables", "Consumer durables", "Manufacturing", "Energy", "Business equipment", "Telecommunications", "Shops", "Health", "Utilities", and "Others". Let $w_1, \ldots, w_{100}$ be the portfolio weights. If the first 10 stocks are labeled as "Consumer Non-durables", the second 10 stock are in "Consumer durables", etc, write down the constraints of the portfolio. 

1. the "health stocks" are no more than 15\% and "energy stocks" are no more than 30\%. 

2. no exposure to Telecommunications. 

3. exposure to "Consumer durables" but gross exposure is zero. 

\newpage

# 7.12 

Let the study period be Jan 2001 to Jan 2015. Apply the sample covariance matrix, the FF 3 factor model, and the RiskMetrics with $\lambda = .94$ to obtain the time varying covariance matrix for Dell, Ford, GE, IBM, Intel, J&J, Merck, 3-mo Tres, and S&P500 at the beginning of each month. Optimize the portfolio and holds for the next 21 days. Compute the risk of such a portfolio and compare it with the equally weighted portfolio. 

First, we get the data we need: 

```{r, eval = FALSE, message = FALSE, warning = FALSE}
start <- as.Date("2001-01-01")
end <- as.Date("2015-01-01")
assets <- c(# "DVMT" # DELL is DROPPED since not provided by Quantmod
  "F", 
  "GE", 
  "IBM", 
  "INTC",
  "JNJ",
  "MRK",
  "SPY")
getSymbols(assets, from = start, to = end) 
# get the 3 mo tresury data 
getSymbols("DGS3MO", src = "FRED")
# combine all info
closing.prices <- merge.xts(DGS3MO, 
                            F[, 4], 
                            GE[, 4],
                            IBM[, 4],
                            INTC[, 4],
                            JNJ[, 4], 
                            MRK[, 4], 
                            SPY[, 4])
# filter out to only dates of interest
data <- closing.prices["2001-01-01/2015-01-01"]
# save
saveRDS(data, "~/workspace/st790-financial-stats/hw5/covdata.rds")
```

Now we can reorganize the data within our desired time frame: 

```{r, eval = FALSE, message = FALSE, warning = FALSE}
data <- readRDS("~/workspace/st790-financial-stats/hw5/covdata.rds")

# break it down monthly 
by_month <- data.frame() 
for(i in 1:ncol(data)){
  temp <- data[, i]
  monthly <- monthlyReturn(temp)
  by_month <- cbind(by_month, monthly)
  colnames(by_month)[i] <- strsplit(names(temp), "[.]")[[1]][1]
}
by_month <- na.omit(by_month)
# get the Fama French data 
FF <- matrix(scan("~/workspace/st790-financial-stats/hw5/F-F_Research_Data_Factors_daily.txt", skip = 5), ncol = 5, byrow = T)
# match dates 
D <- time(by_month)
D <- paste0(substr(D, 1, 4),
            substr(D, 6, 7), 
            substr(D, 9, 10))
ind <- rep(0, length(D))
for(i in 1:length(ind)){
  ind[i] = (1:dim(FF)[1])[D[i] == FF[,1]]
}
ff <- FF[ind, ]
# break it down daily  
dat <- na.omit(data)
by_day <- data.frame() 
for(i in 1:ncol(dat)){
  temp <- dat[, i]
  daily <- dailyReturn(temp)
  by_day <- cbind(by_day, daily)
  colnames(by_day)[i] <- strsplit(names(temp), "[.]")[[1]][1]
}
by_day <- na.omit(by_day)
# get the FF by day 
days <- time(by_day)
days <- paste0(substr(days, 1, 4),
            substr(days, 6, 7), 
            substr(days, 9, 10))
ind2 <- rep(0, length(days))
for(i in 1:length(ind2)){
  ind2[i] = (1:dim(FF)[1])[days[i] == FF[,1]]
}
ff_days <- FF[ind2, ]
saveRDS(by_day, "~/workspace/st790-financial-stats/hw5/byday.rds")
saveRDS(ff_days, "~/workspace/st790-financial-stats/hw5/ff_byday.rds")
saveRDS(by_month, "~/workspace/st790-financial-stats/hw5/bymonth.rds")
saveRDS(ff, "~/workspace/st790-financial-stats/hw5/ff_bymonth.rds")
```

With our data in the desired format, we can now calculate our rolling covariances: 

```{r, message = FALSE, warning = FALSE}
by_day <- readRDS("~/workspace/st790-financial-stats/hw5/byday.rds") 
ff <- readRDS("~/workspace/st790-financial-stats/hw5/ff_byday.rds")
by_day[is.infinite(by_day)] <- 0
by_month <- readRDS("~/workspace/st790-financial-stats/hw5/bymonth.rds")
# risk metrics 
n <- nrow(by_day) 
lambda <- .94 # for daily 
Sigma <- array(0, c(8, 8, n+1))
for(i in 1:n){
  tmp <- as.numeric(c(as.numeric(by_day[i, ])))
  Sigma[,,i+1] <- lambda*Sigma[,,i] + (1-lambda)*tmp %*% t(tmp)
}
# only need optimize monthly 
dates_of_interest <- time(by_month)
# start window in 2002 
dates_of_interest <- dates_of_interest[12:length(dates_of_interest)]
r1 <- r2 <- r3 <- r4 <- r5 <- r6 <- numeric(length(dates_of_interest))
equal_weight <- rep(1, 8)/8
for(i in 1:length(dates_of_interest)){
  temp_date <- dates_of_interest[i]
  # find the correct index to get 252 rolling window 
  temp_D <- paste0(substr(temp_date, 1, 4),
            substr(temp_date, 6, 7), 
            substr(temp_date, 9, 10))
  temp_i <- (1:dim(ff)[1])[temp_D == ff[,1]]
  X <- ff[(temp_i - 252):temp_i, 2:4]
  Y <- by_day[(temp_i - 252):temp_i, ]
  # method 1: sample covariance 
  SigmaS <- cov(Y)
  # method 2: fama french  
  resid <- resid(lsfit(X, Y))
  N <- dim(Y)[1] # length of window 
  SigmaE <- t(resid) %*% resid / (N-4)
  # method 3: risk metrics 
  SigmaR <- Sigma[,,temp_i]
  # optimize  
  result1 <- optimalPortfolio(SigmaS, control = list(type = "minvol"))
  result2 <- optimalPortfolio(SigmaE, control = list(type = "minvol"))
  result3 <- optimalPortfolio(SigmaR, control = list(type = "minvol"))
  # compute risk 
  r1[i] <- sqrt(result1 %*% SigmaS %*% result1) 
  r2[i] <- sqrt(result2 %*% SigmaE %*% result2) 
  r3[i] <- sqrt(result3 %*% SigmaR %*% result3) 
  # compare to equal weight 
  r4[i] <- sqrt(equal_weight %*% SigmaS %*% equal_weight)
  r5[i] <- sqrt(equal_weight %*% SigmaE %*% equal_weight)
  r6[i] <- sqrt(equal_weight %*% SigmaR %*% equal_weight)
}
# viz 
plot(r3, type = "l", ylim = c(0, .03), xlab = "Time", ylabe = "Risk", 
     lty = 1)
lines(r2, col = "red", lty =2 )
lines(r1, col = "blue", lty=3)
legend(10, .025, legend = c("Risk Metrics", 
                           "Fama French", 
                           "Sample Covariance"), 
       col=c("black", "red", "blue"),
       lty=1:3)
title("optimal weights")
# viz 
plot(r6, type = "l", ylim = c(0, .2), xlab = "Time", ylabe = "Risk", 
     lty = 1)
lines(r5, col = "red", lty =2 )
lines(r4, col = "blue", lty=3)
legend(5, .15, legend = c("Risk Metrics", 
                           "Fama French", 
                           "Sample Covariance"), 
       col=c("black", "red", "blue"),
       lty=1:3)
title("equal weights")

```

